### 2016-05-24 dissertacao/agradecimentos.tex

%!TEX root = qualificacao.tex
% RESUMO
\newpage
\thispagestyle{plain}
%\pagenumbering{roman}
\large
\textbf{AGRADECIMENTOS}

%\renewcommand{\baselinestretch}{0.6666666}
Aqui, se tudo der certo, agradecerei um monte de gente.