### 2016-06-15 dissertacao/agradecimentos.tex

%!TEX root = dissertacao.tex
% RESUMO
\newpage
\pagestyle{empty}
\large
\textbf{AGRADECIMENTOS}

%\renewcommand{\baselinestretch}{0.6666666}
Aqui, se tudo der certo, agradecerei um monte de gente.